### 2015-04-15 Paper1Editted/paper1Beditted.tex
Not very clear on what I'm committing

% PLoS doesn't support/use the next package, but we get issues if we don't use this to generate the PDF. Comment this line out when submitting.
\usepackage[section]{placeins}

@@

It is important to distinguish intended action from \textit{motivation}\cite{mcclelland1985motives, rawolle2013relationships}, \textit{cost}\cite{todorov2002optimal, flash1985coordination}, or \textit{goal selection decisions}\cite{ziebart2010modeling}. In typical motor control studies, subjects are motivated to complete an experiment in a timely fashion, and are usually explicitly provided targets to reach to. Classification of intent is prevalent in both lower \cite{strausser2011development, hargrove2013robotic} and upper limb\cite{englehart2003robust, young2012improving} prosthetics where hybrid control algorithms select from among a set of discrete actions (walking/standing/flexion/extension/etc). While subjects may be \textit{motivated} to complete experiments with minimal effort/cost and their \textit{goal} may be to reach a target, here we use intent to describe the course of action (i.e., the trajectory of the arm) taken in service of goals and motives and not the goals nor the motives themselves. Particularly of interest is the intended course of action (i.e, the intended trajectory), even when the actual movement is disturbed and hence no longer matches the intent.

Attempts to deduce motor intent in the past have focused on the assumed spring-like properties of human muscles. Springs produce a force according to their impedance and stretch. By measuring force, impedance, and position, Gomi and Kawato\cite{gomi1997human} were able to deduce stretch and thereby infer the muscle's equilibrium point. Supporters of the ``$\lambda$ model''\cite{feldman1995origin} hypothesized that this equilibrium point represented the intent of a movement even though it did not compensate for the dynamics of the arm as a feedforward controller would. Upon Gomi and Kawato's inspection of the equilibrium point as it evolved in time, it was clear that it was highly complex and often not anatomically realizable. Therefore, it could not well-represent the intent of a simple reaching movement. Unlike the equilibrium point of muscle, the equilibrium point of the entire arm is the path the arm will follow in the absence of disturbance. In the presence of disturbance, the arm might be deflected from its equilibrium. It is presently unknown if or how this arm equilibrium point might change due to disturbances.

@@

\subsection*{Intent Extraction: Deducing the Desired Trajectory}
We first demonstrate the \textit{intent extraction} approach in human motor control, but we later show that the process is applicable to any controlled process with an invertible highest order plant term. The process begins by presuming a model of the controller. Here, we choose the well-known motion control structure of Shadmehr and Mussa-Ivaldi\cite{shadmehr1994adaptive} where the feedforward aspect of the controller perfectly predicts the plant and linearizes the system through cancellation. Additionally, linear feedback rejects position and velocity error. This model was chosen to help illuminate the approach, but later we present how to generalize this approach to more complex models without loss of generality. The equation governing the passive planar dynamics (plant) of musculoskeletal structure is of the form,
\begin{equation}
\underbrace{\overbrace{M(q)\ddot{q}}^{\text{Inertia}}+\overbrace{G(q,\dot{q})}^{\text{Coriolis, Centripal}}}_{\text{Plant}}+E=0
\end{equation}
where $M$ is the mass matrix, $q$ is the joint angles, $\dot{q}$ is joint angular velocity, $\ddot{q}$ is joint angular acceleration, $G$ contains both Coriolis and centripetal effects, and $E$ is any externally-applied torque. The motion behavior changes with the addition of feedforward and/or feedback controllers,
\begin{equation}
\underbrace{\overbrace{M(q)\ddot{q}}^{\text{Inertia}}+\overbrace{G(q,\dot{q})}^{\text{Coriolis, Centripal}}}_{\text{Plant}}+E=\underbrace{\overbrace{\hat{M}(q_d)\ddot{q}_d}^{\text{Inertia}}+\overbrace{\hat{G}(q_d,\dot{q}_d)}^{\text{Coriolis, Centripal}}+\hat{E}}_{\text{Feedforward Controller}}+\underbrace{K_p(q_d-q)+K_d(\dot{q}_d-\dot{q})}_{\text{Impendance, Feedback Controller}}
\end{equation}
where terms with hats over them ($\hat{M}$, $\hat{G}$, or $\hat{E}$) indicate that they represent the nervous system's best estimate of the forces and dynamics it will encounter, which is also known as an internal model \cite{shadmehr1994adaptive}. This portion of the system serves as an inverse-dynamics feedforward controller that cancels out the dynamics of the arm in the torque balance. If the nervous system has sufficient experience and is expecting $E$, it is included as part of the internal model, $\hat{E}$, otherwise $\hat{E}$ is set to zero. $K_p$ and $K_d$ are the lumped impedance and feedback terms that employ a moving state equilibrium to accomplish the desired trajectory, $q_d$. This $q_d$ has a dual meaning in that it signifies both the unknown desired trajectory that we seek to discover and also the moving equilibrium trajectory of the arm.

@@

To explore the conditions under which this procedure for recovering the intent trajectory can be generalized, we must take a series of counter-intuitive steps. First, we distinguish between joint space and generalized space by introducing a generalized state variable, $x$. Second, we separate the plant (arm) into its process (dynamics), $P$, and actuator (muscles), $A$, components. Third, we model these components in a very general sense as time-varying linear systems,
\begin{equation} \label{eq:generalPlant}
\underbrace{\overbrace{\sum_{n=1}^N P_nx^{(n)}}^\text{Process}+\overbrace{\sum_{m=1}^M A_m(x-x_e)^{(m)}}^\text{Actuator}}_\text{Plant}+E=0
\end{equation}
where $x_e$ is the equilibrium trajectory of the actuator. Linearization is informative here because stable equilibrium points of linearizations of nonlinear systems are also equilibrium points of the nonlinear system\cite[p.~156]{khalil2002nonlinear} as long as the Jacobian is bounded and Lipschitz continuous\cite[p.~161]{khalil2002nonlinear}. Fourth, we describe the actuator control law using the same expansions,
\begin{equation}
\overbrace{\sum_{n=1}^N \hat{P}_n x^{(n)}_d+\hat{E}}^\text{Internal Model}+\overbrace{\sum_{m=1}^M A_m(x_d-x_e)^{(m)}}^\text{Actuator}=0
\end{equation}
where an internal model, $\hat{P}$ and $\hat{E}$, predicts system dynamics and external disturbances in order to determine the actuator equilibrium $x_e$ such that $x$ will track a desired path $x_d$. The feedforward component must use the actuator and hence it shares the actuator's equilibrium with the plant. This allows us to finally relate the physical system and its control law by solving for $x_e$ and substituting into \eqref{eq:generalPlant},
\begin{equation}
\overbrace{\sum_{n=1}^N P_nx^{(n)}}^\text{Process}+E=\overbrace{\sum_{n=1}^N \hat{P}_n x^{(n)}_d+\hat{E}}^\text{Internal Model}+\overbrace{\sum_{m=1}^M A_m (x_d-x)^{(m)}}^\text{Feedback (Actuator)}
\end{equation}
wherein $x_e$ vanishes recovering our familiar model. While a proper choice of $x_e$ is perhaps essential for control, our approach does not require a model relating this equilibrium and intended trajectory. Note also that the actuator's equilibrium $x_e$ is not the process's equilibrium unless all derivatives of $x_d$ are zero. This is important as the actuator state ($\lambda$) has been claimed to be the intended movement\cite{feldman1995origin}. This upholds Gomi and Kawato's finding that the arm muscles' equilibrium does not represent reaching intent\cite{gomi1997human}. If the highest order coefficient of the dynamic model of the process $\hat{P}_N$ can be inverted, the impedances can be modeled, and the linearized system is stable, it is possible to solve for $x_d^{(N)}$ and integrate for $x_d$ revealing the intended trajectory.

@@

\subsection*{Experimental Design}
We chose 15 centimeters long simulated reaches of the right arm, beginning at 38 centimeters out from and 5.7 centimeters left of the right shoulder and ending at 38 centimeters out from and 10.7 centimeters right of the right shoulder.  We used the following two types of perturbing forces for each combination of distance and direction: 
\begin{enumerate}
\item Pulse forces applied in one of the the two directions perpendicular to the direction of movement began when the subject had moved either 10\% or 50\% of the distance to the target and lasted for 150ms.
\item Noise forces began once the subject moved 3 millimeters, and lasted for the duration of the motion. The forces were drawn from a white noise generator at 1000 Hz with flat power spectral density of 1N, and then passed through a 4th order low-pass Butterworth filter with cutoff $10 \pi$ rad/s.
\end{enumerate}
Our error metric was deviation in position, where unsigned error was calculated as the mean magnitude of deviation from the straight-line nominal trajectory to the target. This was evaluated across each of 150 bins measuring 1 millimeter in width and spaced evenly throughout the 15 centimeter reach. Mean unsigned error (MUE) was also used to summarize the overall error in each movement for sensitivity analysis. Mean and maximum signed error relative to the direction of disturbance, also called perpendicular deviation, were used to measure reaching accuracy in order to elucidate direction of any corrections.

@@

\subsection*{Statistical Analysis}
Our principal goal was to determine if and when disturbed trajectories departed from undisturbed trajectories. To detect this departure, we use the one-tailed Student's t-test $(\alpha=.05)$ at 5 millisecond intervals to determine whether or not the deviation of the disturbed trajectories had exceeded the maximum deviations of the undisturbed trajectories. The time span between departure of the hand and estimated intent was compared against 120 milliseconds using the nonparametric \textit{sign test}. The MATLAB statistics toolbox package \cite{MATLAB:2008} was used for all comparisons.

@@

The words \textit{intent} and \textit{equilibrium} are contentious because of the various definitions that exist in the literature. Our simple definition here is that the intent is the path that would have been taken had there been no external disturbances. A rich history of research has suggested many ways that an equilibrium trajectory might explain how the central nervous system \cite{feldman1995origin} and/or pre-planned elements \cite{shadmehr1994adaptive} can explain action. Here we suggest that one should directly define intent in the context of a hypothesized control model. Our work here assumed the model of Burdet \cite{burdet2006stability}. However, other, more sophisticated, models may be deemed more appropriate in the future. What is clear is that given some model of control, this intent extraction approach is able to invert the process to estimate intent.

@@

%\section*{References}
\bibliography{paper1.bib}
% Either type in your references using
% \begin{thebibliography}{}
% \bibitem{}
% Text
% \end{thebibliography}
%
% OR
%
% Compile your BiBTeX database using our plos2009.bst
% style file and paste the contents of your .bbl file
% here.
% 
\pagebreak
\section*{Figure Legends}
% This section is for figure legends only, do not include
% graphics in your manuscript file.
%
%\begin{figure}
%\caption{
%{\bf Bold the first sentence.}  Rest of figure caption.  
%}
%\label{Figure_label}
%\end{figure}
\begin{figure}
\caption{
{\bf Simulated data illustrating tautology of extraction across pulse and filtered Gaussian noise disturbance types.} Intent is modeled as a minimum jerk, 5th order polynomial. Forces experienced are combined with intent via Burdet et al.'s \cite{burdet2006stability} model to produce the simulated arm trajectory. Extraction to recover intention from arm and force trajectory follows. Parameter errors are introduced into the extraction and varied according to Table 1 to estimate sensitivity. Examining the distribution of deviation in trajectories from this analysis reveals that sensitivity is not uniformly distributed throughout a reach.
}
\label{fig:synthetic}
\end{figure}

@@

\begin{figure}
\caption{
{\bf Subjects' disturbed hand trajectories and extraction of intended trajectories from them reveal that both variance and invariance of the motor plan can occur even in response to very large disturbances.} The hand path, in black, deviates from the blue baseline, as a force pulse, gray arrows, is applied. Three of the eight subjects' intent trajectories (orange) did not significantly deviate from undisturbed movement (panel A) as observed by the intent (orange) region (mean $\pm$ 95\% confidence interval on the mean) not departing the undisturbed movement (blue) region. The time at which the hand (black) region departed from undisturbed movement is marked with a green dot for each subject. Five of the eight subjects' intent trajectories did significantly deviate from undisturbed movement (panel B). The time of intent departure from the undisturbed trajectories is marked with a red dot for each subject.
}
\label{fig:anecdotes}
\end{figure}

\pagebreak

>>>>

\usepackage[section]{placeins}

@@

It is important to distinguish intended action from \textit{motivation}\cite{mcclelland1985motives, rawolle2013relationships}, \textit{cost}\cite{todorov2002optimal, flash1985coordination}, or \textit{goal selection decisions}\cite{ziebart2010modeling}. In typical motor control studies, subjects are motivated to complete an experiment in a timely fashion, and are usually explicitly provided targets to reach to. Classification of intent is prevalent in both lower \cite{strausser2011development, hargrove2013robotic} and upper limb\cite{englehart2003robust, young2012improving} prosthetics where hybrid control algorithms select from among a set of discrete actions (walking/standing/flexion/extension/etc). While subjects may be \textit{motivated} to complete experiments with minimal effort/cost and their \textit{goal} may be to reach a target, here we use intent to describe the course of action (i.e., the trajectory of the arm) taken in service of goals and motives and not the goals nor the motives themselves. Particularly of interest is the intended course of action (i.e, the intended trajectory), even when the actual movement is disturbed and hence no longer matches the intent. In other words, we operationally define \textit{intent} as a subjunctive -- where would the motion have gone had it not been disturbed?

Attempts to deduce motor intent in the past have focused on the assumed spring-like properties of human muscles. Springs produce a force according to their impedance and stretch. By measuring force, impedance, and position, Gomi and Kawato\cite{gomi1997human} were able to deduce stretch and thereby infer the muscle's equilibrium point. Supporters of the ``$\lambda$ model''\cite{feldman1995origin} hypothesized that this muscle equilibrium point, and not the equilibrium of the whole arm, represented the intent of a movement even though it did not compensate for the dynamics of the arm as a feedforward controller would. Upon Gomi and Kawato's inspection of the muscle equilibrium point as it evolved in time, it was clear that it was highly complex and often not anatomically realizable. Therefore, it could not well-represent the intent of a simple reaching movement. Unlike the equilibrium point of muscle, the equilibrium point of the entire arm is the path the arm will follow in the absence of disturbance. In the presence of disturbance, the arm might be deflected from its equilibrium. It is presently unknown if or how this arm equilibrium point might change due to disturbances.

@@

\subsection*{Intent Extraction: Deducing the Desired Trajectory}
We first demonstrate the \textit{intent extraction} approach in human motor control, but we later show that the process is applicable to any controlled process with an invertible highest order plant term. The process begins by presuming a model of the controller. Here, we choose the well-known motion control structure of Shadmehr and Mussa-Ivaldi\cite{shadmehr1994adaptive} where the feedforward aspect of the controller perfectly predicts the plant and linearizes the system through cancellation. Additionally, linear feedback rejects position and velocity error. This model was chosen to help illuminate the approach, but later we present how to generalize this approach to more complex models without loss of generality. The equation governing the passive planar dynamics (plant) of musculoskeletal structure is of the form,
\begin{equation}
\underbrace{\overbrace{M(q)\ddot{q}}^{\text{Inertia}}+\overbrace{G(q,\dot{q})}^{\text{Coriolis, Centripal}}}_{\text{Plant}}+E=0
\end{equation}
where $M$ is the mass matrix, $q$ is the joint angles, $\dot{q}$ is joint angular velocity, $\ddot{q}$ is joint angular acceleration, $G$ contains both Coriolis and centripetal effects, and $E$ is any externally-applied torque. The motion behavior changes with the addition of feedforward and/or feedback controllers,
\begin{equation}
\underbrace{\overbrace{M(q)\ddot{q}}^{\text{Inertia}}+\overbrace{G(q,\dot{q})}^{\text{Coriolis, Centripal}}}_{\text{Plant}}+E=\underbrace{\overbrace{\hat{M}(q_d)\ddot{q}_d}^{\text{Inertia}}+\overbrace{\hat{G}(q_d,\dot{q}_d)}^{\text{Coriolis, Centripal}}+\hat{E}}_{\text{Feedforward Controller}}+\underbrace{K_p(q_d-q)+K_d(\dot{q}_d-\dot{q})}_{\text{Impendance, Feedback Controller}}
\end{equation}
where terms with hats over them ($\hat{M}$, $\hat{G}$, or $\hat{E}$) indicate that they represent the nervous system's best estimate of the forces and dynamics it will encounter, which is also known as an internal model \cite{shadmehr1994adaptive}. This portion of the system serves as an inverse-dynamics feedforward controller that cancels out the dynamics of the arm in the torque balance. If the nervous system has sufficient experience and is expecting $E$, it is included as part of the internal model, $\hat{E}$; otherwise $\hat{E}$ is set to zero. $K_p$ and $K_d$ are the lumped impedance and feedback terms that employ a moving state equilibrium to accomplish the desired trajectory, $q_d$. This $q_d$ has a dual meaning in that it signifies both the unknown desired trajectory that we seek to discover and also the moving equilibrium trajectory of the arm.

@@

To explore the conditions under which this procedure for recovering the intent trajectory can be generalized, we take a series of steps to leverage a relationship between the equilibria of a nonlinear system and the equilibria of linearizations of that system. First, we distinguish between joint space and generalized space by introducing a generalized state variable, $x$. Second, we separate the plant (arm) into its process (dynamics), $P$, and actuator (muscles), $A$, components. Third, we model these components in a very general sense as time-varying linear systems,
\begin{equation} \label{eq:generalPlant}
\underbrace{\overbrace{\sum_{n=1}^N P_nx^{(n)}}^\text{Process}+\overbrace{\sum_{m=1}^M A_m(x-x_e)^{(m)}}^\text{Actuator}}_\text{Plant}+E=0
\end{equation}
where $x_e$ is the equilibrium trajectory of the actuator. Linearization is informative here because stable equilibrium points of linearizations of nonlinear systems are also equilibrium points of the nonlinear system\cite[p.~156]{khalil2002nonlinear} as long as the Jacobian is bounded and has a bounded time derivative (Lipschitz continuous)\cite[p.~161]{khalil2002nonlinear}. Fourth, we describe a  control law for the actuator using the same expansions,
\begin{equation}
\overbrace{\sum_{n=1}^N \hat{P}_n x^{(n)}_d+\hat{E}}^\text{Internal Model}+\overbrace{\sum_{m=1}^M A_m(x_d-x_e)^{(m)}}^\text{Actuator}=0
\end{equation}
where an internal model, $\hat{P}$ and $\hat{E}$, predicts system dynamics and external disturbances in order to determine the actuator equilibrium $x_e$ such that $x$ will track a desired path $x_d$. The feedforward component must use the actuator and hence it shares the actuator's equilibrium with the plant. This allows us to finally relate the physical system and its control law by solving for $x_e$ and substituting into \eqref{eq:generalPlant},
\begin{equation}
\overbrace{\sum_{n=1}^N P_nx^{(n)}}^\text{Process}+E=\overbrace{\sum_{n=1}^N \hat{P}_n x^{(n)}_d+\hat{E}}^\text{Internal Model}+\overbrace{\sum_{m=1}^M A_m (x_d-x)^{(m)}}^\text{Feedback (Actuator)}
\end{equation}
wherein $x_e$ vanishes recovering our familiar model. While a proper choice of $x_e$ is perhaps essential for control, our approach does not require a model relating this equilibrium and intended trajectory. Note also that the actuator's equilibrium $x_e$ is not the process's equilibrium unless all derivatives of $x_d$ are zero. This is important as the actuator state ($\lambda$) has been claimed to be the intended movement\cite{feldman1995origin}. This upholds Gomi and Kawato's finding that the arm muscles' equilibrium does not represent reaching intent\cite{gomi1997human}. If the highest order coefficient of the dynamic model of the process $\hat{P}_N$ can be inverted, the impedances can be modeled, and the linearized system is stable, it is possible to solve for $x_d^{(N)}$ and integrate for $x_d$, revealing the intended trajectory.

@@

\subsection*{Experimental Design}
We chose 15 centimeter long simulated reaches of the right arm, beginning at 38 centimeters out from and 5.7 centimeters left of the right shoulder and ending at 38 centimeters out from and 10.7 centimeters right of the right shoulder.  We used the following two types of perturbing forces for each combination of distance and direction: 
\begin{enumerate}
\item Pulse forces applied in one of the the two directions perpendicular to the direction of movement began when the subject had moved either 10\% or 50\% of the distance to the target and lasted for 150ms.
\item Noise forces began once the subject moved 3 millimeters, and lasted for the duration of the motion. The forces were drawn from a white noise generator at 1000 Hz with flat power spectral density of 1N, and then passed through a 4th order low-pass Butterworth filter with cutoff $10 \pi$ rad/s.
\end{enumerate}
Our error metric was deviation in position, where unsigned error was calculated as the mean magnitude of deviation from the straight-line nominal trajectory to the target. This was evaluated across each of 150 bins measuring 1 millimeter in width and spaced evenly throughout the 15 centimeter reach. Mean unsigned error (MUE) was also used to summarize the overall error in each movement for sensitivity analysis. Mean and maximum signed error relative to the direction of disturbance, also called perpendicular deviation, were used to measure reaching accuracy in order to elucidate direction of any corrections.

@@

\subsection*{Statistical Analysis}
Our principal goal was to determine if and when disturbed trajectories departed from undisturbed trajectories. To detect this departure, we use the one-tailed Student's t-Test $(\alpha=.05)$ at 5 millisecond intervals to determine whether or not the deviation of the disturbed trajectories had exceeded the maximum deviations of the undisturbed trajectories. The median time span between departure of the hand and departure of the estimated intent was compared against 120 milliseconds using the nonparametric Sign Test. The MATLAB statistics toolbox package \cite{MATLAB:2008} was used for all comparisons.

@@

The words \textit{intent} and \textit{equilibrium} are contentious because of the various definitions that exist in the literature. Our simple definition here is that the intent is the path that would have been taken had there been no external disturbances. A rich history of research has suggested many ways that an equilibrium trajectory can explain action through feedback responses \cite{feldman1995origin} and/or pre-planned elements \cite{shadmehr1994adaptive}. Here we suggest that one should directly define intent in the context of a hypothesized control model. Our work here assumed the model of Burdet \cite{burdet2006stability}. However, other, more sophisticated, models may be deemed more appropriate in the future. What is clear is that given some model of control, this intent extraction approach is able to invert the process to estimate intent.

@@

%\section*{References}
\bibliography{paper1.bib}
% Either type in your references using
% \begin{thebibliography}{}
% \bibitem{}
% Text
% \end{thebibliography}
%
% OR
%
% Compile your BiBTeX database using our plos2009.bst
% style file and paste the contents of your .bbl file
% here.
%
\pagebreak

\section*{Figure Legends}
% This section is for figure legends only, do not include
% graphics in your manuscript file.
%
%\begin{figure}
%\caption{
%{\bf Bold the first sentence.}  Rest of figure caption.  
%}
%\label{Figure_label}
%\end{figure}
\begin{figure}
\caption{
{\bf Simulated data illustrating tautology of extraction across pulse and filtered Gaussian noise disturbance types.} Intent is modeled as a minimum jerk, 5th order polynomial. Forces experienced are combined with intent via Burdet et al.'s \cite{burdet2006stability} model to produce the simulated arm trajectory. Extraction to recover intention from arm and force trajectory follows. Parameter errors are introduced into the extraction and varied according to Table 1 to estimate sensitivity. Examining the distribution of deviation in trajectories from this analysis reveals that sensitivity is not uniformly distributed throughout a reach.
}
\label{fig:synthetic}
\end{figure}

@@

\begin{figure}
\caption{
{\bf Subjects' disturbed hand trajectories and extraction of intended trajectories from them reveal that both variance and invariance of the motor plan can occur even in response to very large disturbances.} The hand path, in black, deviates from the blue baseline, as a force pulse, gray arrows, is applied. Three of the eight subjects' intent trajectories (orange) did not significantly deviate from undisturbed movement (panel A) as observed by the intent region not departing the undisturbed movement (blue) region. Shaded regions represent mean $\pm$ 95\% confidence interval on the mean. The first time at which the hand (black) region departed from undisturbed movement is marked with a green dot for each subject. Five of the eight subjects' intent trajectories did significantly deviate from undisturbed movement (panel B). The time of intent departure from the undisturbed trajectories is marked with a red dot for each subject.
}
\label{fig:anecdotes}
\end{figure}
\pagebreak
